\thetitle

### 2021-03-02 writing/why-rational.tex
landscape, why-rational: fix mistakes in my previous changes
adding methodology subsection

This paper dubs Lazarek et. all's idea {\em the rational programmer\/}. Like Stuart
Mill's {\it homo economicus\/}, the rational programmer approximates the
behavior of a software developer who reduces time spent on a task given the
available information. In the context of gradual typing, the rational programmer
has two pieces of information when an impedance mismatch signals exceptional
behavior: the error message and the state of the program. Hence, the most
rational procedure is to use the former to improve the latter. Specifically,
a rational programmer translates the Wadler-Findler slogan into a debugging
method, searching for the source of the impedance mismatch by adding type
annotations to the parts of the program identified in the error.
If the type checker rejects an annotation, the rational programmer has found the problem.
Measuring this simulated programmer's behavior on a large number of
debugging scenarios yields data that is similar to data collected in an
empirical, human-facing manner.

@@

Assuming the rational programmer trusts the type of {\tt na-client}, the next
step is to inspect the {\tt layer} module and to equip it with type
annotations. This results in the \typecolor\ module in the middle, and {\tt
main}'s import is now re-directed there by {\tt (submod ".." layer)}. Running the
program again yields a new message:
\begin{verbatim}
    neg-abs: broke its own contract
      promised: (<=/c 0)
      produced: 10
      in: (-> any/c (<=/c 0))
      contract from: (interface for neg-abs)
      blaming: (interface for neg-abs)
       (assuming the contract is correct)
\end{verbatim}
The last step of the rational programmer is to assign types to the {\tt server}
module. At this point the type checker objects to the conjecture type of {\tt
neg-abs}---and the source of the impedance mismatch is found. How to fix it is a
separate question.

@@

>>>>

This paper dubs Lazarek et. al's idea {\em the rational programmer\/}. Like Stuart
Mill's {\it homo economicus\/}, the rational programmer approximates the
behavior of a software developer who reduces time spent on a task given the
available information. In the context of gradual typing, the rational programmer
has two pieces of information when an impedance mismatch signals exceptional
behavior: the error message and the state of the program. Hence, the most
rational procedure is to use the former to improve the latter. Specifically,
a rational programmer translates the Wadler-Findler slogan into a debugging
method, searching for the source of the impedance mismatch by adding type
annotations to the parts of the program identified in the error.
If the type checker rejects an annotation, the rational programmer has found the problem.
Measuring this simulated programmer's behavior on a large number of
debugging scenarios yields data that is similar to data collected in an
empirical, human-facing manner.

@@

Assuming the rational programmer trusts the type of {\tt na-client}, the next
step is to inspect the {\tt layer} module and to equip it with type
annotations. The result is the \typecolor\ module in the middle, and {\tt
main}'s import is now re-directed there by {\tt (submod ".." layer)}. Running the
program again yields a new message:
\begin{verbatim}
    neg-abs: broke its own contract
      promised: (<=/c 0)
      produced: 10
      in: (-> any/c (<=/c 0))
      contract from: (interface for neg-abs)
      blaming: (interface for neg-abs)
       (assuming the contract is correct)
\end{verbatim}
The last step of the rational programmer is to assign types to the {\tt server}
module. At this point the type checker objects to the conjecture type of {\tt
neg-abs}---and the source of the impedance mismatch is found. How to fix it is a
separate question.

@@

\subsection{How to Turn the Idea of the Rational Programmer to a Methodology}

\input{rational-methodology.tex}
